\LARGE \bf Steering a Swarm Using Global Inputs\\ and Swarm Statistics

### 2017-03-27 journalSwarmControl/expResults.tex
shortened the paper to 12 pages.
Worked on related work and cut some more of the paper

Transport of goods and materials between points is at the heart of all engineering and construction in real-world systems. This experiment varied from 1 to 500 the population of particles used to transport an object. The total area, maximum particle speed, and total net force the swarm could produce were constant. The particles pushed a large hexagonal object through an  {\sffamily S}-shaped maze. We hypothesized participants would complete the task faster with more particles. The results, shown in Fig.~\ref{fig:ResVaryNu}, do not support our hypothesis, indicating a minimum around 130 particles.

@@

Sensing is expensive, especially on the nanoscale. To see nanocars~\cite{Chiang2011} fastened molecules that fluoresce light when activated by a strong light source. Unfortunately, multiple exposures can destroy these molecules, a process called \emph{photobleaching}. Photobleaching can be minimized by lowering the excitation light intensity, but \cite{Cazes2001} showed this increases the probability of missed detections.  This experiment explores manipulation with varying amounts of sensing information: {\bf full-state} sensing provides the most information by showing the position of all particles; {\bf convex-hull} draws a convex hull around the outermost particles; {\bf mean} provides the average position of the population; and {\bf mean + variance} adds a confidence ellipse. Fig.~\ref{fig:Visualization} shows screenshots of the same particle swarm with each type of visual feedback. Full-state requires $2n$ data points for $n$ particles. Convex-hull requires at worst $2n$, but according to \cite{har2011expected}, the expected number is $O(2 n^{1/3})$.  Mean requires two, and variance three, data points.  Mean and mean + variance are convenient even with millions of particles.

@@

\paragraph{Varying noise:}
Micro-particles are affected by random collisions with molecules. The effect of these collisions is called Brownian motion.
This experiment varied the strength of these disturbances to study how noise affects human control of large swarms. Noise was applied at every timestep:
\begin{align*}
\dot{x}_i &= u_x + m_i\cos(\psi_i)\\
 \dot{y}_i &= u_y + m_i\sin(\psi_i).
 \end{align*}
$m_i,\psi_i$ were uniformly IID, with $m_i\in[0,M]$ and $\psi_i\in[0,2\pi]$. $M$ was a constant for each trial ranging from 0 to 200\% of the maximum control power ($u_{max}$).

@@

Time increases for $n>4$ particles. For $n\in[1,4]$ the goal patterns are not hollow, but starting at $n$=5 they are.  A better experiment design would randomly place the goal positions.  Initially we tried this, but beta-testers strongly disliked trying to arrange particles in random patterns.

>>>>

%Transport of goods and materials between points is at the heart of all engineering and construction in real-world systems. 
This experiment varied from 1 to 500 the population of particles used to transport an object. The total area, maximum particle speed, and total net force the swarm could produce were constant. The particles pushed a large hexagonal object through an  {\sffamily S}-shaped maze. We hypothesized participants would complete the task faster with more particles. The results, shown in Fig.~\ref{fig:ResVaryNu}, do not support our hypothesis, indicating a minimum around 130 particles.

@@

%Sensing is expensive, especially on the nanoscale. To see nanocars~\cite{Chiang2011} fastened molecules that fluoresce light when activated by a strong light source. 
%Unfortunately, multiple exposures can destroy these molecules, a process called \emph{photobleaching}. 
%Photobleaching can be minimized by lowering the excitation light intensity, but \cite{Cazes2001} showed this increases the probability of missed detections.  
This experiment explores manipulation with varying amounts of sensing information: {\bf full-state} sensing provides the most information by showing the position of all particles; {\bf convex-hull} draws a convex hull around the outermost particles; {\bf mean} provides the average position of the population; and {\bf mean + variance} adds a confidence ellipse. Fig.~\ref{fig:Visualization} shows screenshots of the same particle swarm with each type of visual feedback. Full-state requires $2n$ data points for $n$ particles. Convex-hull requires at worst $2n$, but according to \cite{har2011expected}, the expected number is $O(2 n^{1/3})$.  Mean requires two, and variance three, data points.  Mean and mean + variance are convenient even with millions of particles.

@@

\paragraph{Varying noise:}
%Micro-particles are affected by random collisions with molecules. The effect of these collisions is called Brownian motion.
This experiment varied the strength of %these 
disturbances to study how noise affects human control of large swarms. Noise was applied at every timestep:
\begin{align*}
\dot{x}_i &= u_x + m_i\cos(\psi_i)\\
 \dot{y}_i &= u_y + m_i\sin(\psi_i).
 \end{align*}
$m_i,\psi_i$ were uniformly IID, with $m_i\in[0,M]$ and $\psi_i\in[0,2\pi]$. $M$ was a constant for each trial ranging from 0 to 200\% of the maximum control power ($u_{max}$).

@@

%Time increases for $n>4$ particles. For $n\in[1,4]$ the goal patterns are not hollow, but starting at $n$=5 they are.  A better experiment design would randomly place the goal positions.  Initially we tried this, but beta-testers strongly disliked trying to arrange particles in random patterns.

### 2017-03-27 journalSwarmControl/relatedWork.tex
Worked on related work and cut some more of the paper

\subsection{Global control of microrobots}
Small robots have been constructed with physical heterogeneity so that they respond differently to a global, broadcast control signal.  Examples include \emph{scratch-drive microrobots}, actuated and controlled by a DC voltage signal from a substrate by \cite{Donald2006,Donald2008};   magnetic structures  with different cross-sections that could be independently steered by \cite{Floyd2011,Diller2013};   \emph{MagMite} microrobots with different resonant frequencies and a global magnetic field by \cite{Frutiger2008}; and  magnetically controlled nanoscale helical screws constructed to stop movement at different cutoff frequencies of a global magnetic field by
\cite{Tottori2012} and \cite{Peyer2013}. 
%In our previous work with robots that can be modeled as nonholonomic unicycles~\cite{Becker2012k}, we showed that an inhomogeneity in turning speed is enough to make even an infinite number of robots controllable with regard to position. 
%While these approaches theoretically scale to large populations, they also require both excellent state estimation and heterogeneous robots.

@@

 \paragraph{Sensing---large populations:}
 Parallel control of $n$ differential-drive robots in a 2D workspace requires $3n$ state variables. Even holonomic robots require $2n$ state variables. Numerous methods exist for measuring this state in microrobotics.  These solutions use computer vision systems to sense position and heading angle, with corresponding challenges of handling missed detections and image registration between detections and robots.  These challenges are increased at the nanoscale where sensing competes with control for communication bandwidth.   We examine control when the operator has access to partial feedback, including only the first and/or second moments of a population's position, or only the convex-hull containing the robots.

@@

@@

Our $n$-robot system with 2 control inputs and 4$n$ states is inherently under-actuated, and superficially bears resemblance to compliant, under-actuated manipulators.  Our swarms conform to the object to be manipulated.  However our swarms lack the restoring force provided by flexures in \cite{odhner2014compliant} or silicone in \cite{deimel2014novel}.  Our swarms tend to disperse, so we require  artificial forces, such as the variance control primitives in \S \ref{sec:VarianceControl},
 to regroup the swarms.

>>>>

\subsection{Global control of microrobots}
Small robots have been constructed with physical heterogeneity so that they respond differently to a global, broadcast control signal.  Examples include \emph{scratch-drive microrobots}, actuated and controlled by a DC voltage signal from a substrate by \cite{Donald2006,Donald2008};   magnetic structures  with different cross-sections that could be independently steered by \cite{Floyd2011,Diller2013};   \emph{MagMite} microrobots with different resonant frequencies and a global magnetic field by \cite{Frutiger2008}; and  magnetically controlled nanoscale helical screws constructed to stop movement at different cutoff frequencies of a global magnetic field by
\cite{Tottori2012} and \cite{Peyer2013}.

%In our previous work with robots that can be modeled as nonholonomic unicycles~\cite{Becker2012k}, we showed that an inhomogeneity in turning speed is enough to make even an infinite number of robots controllable with regard to position. 
%While these approaches theoretically scale to large populations, they also require both excellent state estimation and heterogeneous robots.

@@

 \paragraph{Sensing---large populations:}
$n$ differential-drive robots in a 2D workspace require $3n$ state variables. Even holonomic robots require $2n$ state variables. Numerous methods exist for measuring this state in microrobotics.  These solutions use computer vision systems to sense position and heading angle, with corresponding challenges of handling missed detections and image registration between detections and robots.  These challenges are increased at the nanoscale where sensing competes with control for communication bandwidth.   We examine control when the operator has access to partial feedback, including only the first and/or second moments of a population's position, or only the convex-hull containing the robots.

@@

  Most humans are able, with practice, steer a swarm of robots controlled by a global input. Prior to our paper, no algorithm existed. Using human input to learn how to control a dynamic system is a line of research with a rich history \cite{argall2009, abbeel2010}.This paper exploits insights gained from ``swarmcontrol.net", particularly the fact that having a swarm's mean and variance is sufficient for object manipulation through an obstacle field.

@@

Our $n$-robot system with 2 control inputs and 4$n$ states is inherently under-actuated, and superficially bears resemblance to compliant, under-actuated manipulators.  Our swarms conform to the object to be manipulated, but lack the restoring force provided by flexures in \cite{odhner2014compliant} or silicone in \cite{deimel2014novel}.  Our swarms tend to disperse, so we require  artificial forces, such as the variance control primitives in \S \ref{sec:VarianceControl},
 to regroup the swarms.

### 2017-03-27 journalSwarmControl/theory.tex
Worked on related work and cut some more of the paper

\end{bmatrix} = \begin{bmatrix}
0 & 1 & \ldots & 0 & 0 \\
0 & 0 & \ldots& 0 & 0 \\
\vdots &  \vdots & \ddots & \vdots & \vdots \\
0 & 0  & \ldots & 0 & 1 \\
0 & 0 & \ldots& 0 & 0 
\end{bmatrix}  \begin{bmatrix}
x_{1}\\
\dot{x}_{1}\\
\vdots \\
x_{n}\\
\dot{x}_{n}
\end{bmatrix} + \begin{bmatrix}
0\\
1\\
\vdots\\
0\\
1
\end{bmatrix} u_x
\end{equation}
 However, just as with one robot, we can only control two states because the controllability matrix $\mathcal{C}_n$ has rank two:
\begin{equation}
\mathcal{C}_n=\left[ \begin{matrix} 
0\\
1\\
\vdots\\
0\\
1
\end{matrix}
\,\middle\vert\,
  \begin{matrix} 
1\\
0\\
\vdots\\
1\\
0
\end{matrix}
\,\middle\vert\,
\begin{matrix} 
0\\
0\\
\vdots\\
0\\
0
\end{matrix}
\,\middle\vert\,
\begin{matrix} 
0\\
0\\
\vdots\\
0\\
0
\end{matrix}\,\middle\vert\,
\ldots \right], \quad \textrm{rank}(\mathcal{C}_n)=2.
\end{equation}  
\subsection{Controlling the mean position}\label{sec:controlMeanPosition}
This means \emph{any} number of robots controlled by a global command have just two controllable states in each axis. We cannot arbitrarily control the position and velocity of two or more robots, but have options on which states to control.  One option is to control the position and velocity of the $j^{th}$ robot. To find a potentially more useful option, we create the following reduced order system that represents the mean $x$ position and velocity of the $n$ robots:
\begin{align}
\begin{bmatrix}\nonumber
\dot{\bar{x}} \\
\ddot{\bar{x}}
\end{bmatrix} &= \frac{1}{n} \begin{bmatrix}
0& 1& \ldots &0& 1 \\
0& 0& \ldots &0& 0
\end{bmatrix}
\begin{bmatrix}
x_{1}\\
\dot{x}_{1}\\
\vdots\\
x_{n}\\
\dot{x}_{n}
\end{bmatrix} \\
&+ \frac{1}{n}\begin{bmatrix}
0& 0&  \ldots &0& 0 \\
0& 1&  \ldots &0& 1
\end{bmatrix}\begin{bmatrix} 
0\\
1\\
\vdots\\
0\\
1
\end{bmatrix} u_x
\end{align}
Thus:
\begin{equation}
\begin{bmatrix}
\dot{\bar{x}} \\
\ddot{\bar{x}}
\end{bmatrix} = \begin{bmatrix}
0& 1 \\
0& 0
\end{bmatrix}
\begin{bmatrix}
\bar{x}\\
\dot{\bar{x}}
\end{bmatrix} + \begin{bmatrix} 
0\\
1
\end{bmatrix} u_x
\end{equation}

@@

We will prove the origin is globally asymptotically stabilizable by using a control-Lyapunov function, as in \citet{Lyapunov1992}.  A suitable Lyapunov function is the squared variance error:
\begin{align}
\label{eq:LyapunovVariance}
V(t,\bf{x})  &= \frac{1}{2} (\sigma^2(\mathbf{x}) - \sigma^2_{goal})^2\nonumber\\
\dot{V}(t,\bf{x}) &= (\sigma^2(\mathbf{x})-\sigma^2_{goal})\dot{\sigma}^2(\mathbf{x})
\end{align}
We note here that $V(t,\mathbf{x})$ is positive definite and radially unbounded, and $V(t,\mathbf{x}) \equiv 0$ only at $\sigma^2(\mathbf{x}) = \sigma^2_{goal}$.
To make $\dot{V}(t,\mathbf{x})$ negative semi-definite, we choose
\begin{align}\label{eq:controlVariance}
u(t) &=   \begin{cases}
	 \mbox{move to wall} &\mbox{if } \sigma^2(\mathbf{x})>\sigma^2_{goal} \\ 
	 \mbox{move from wall} & \mbox{if } \sigma^2(\mathbf{x}) \le \sigma^2_{goal}.
\end{cases} 
\end{align}
 For such a $u(t)$,
 \begin{align}
\dot{\sigma}^2(\mathbf{x}) &=   \begin{cases}
	 \mbox{negative} &\mbox{if } \sigma^2(\mathbf{x})> \max(\sigma^2_{goal}, \sigma^2_{optimal}(n,r))  \\ 
	 W \epsilon & \mbox{if } \sigma^2(\mathbf{x}) \le \sigma^2_{goal},
\end{cases} 
\end{align} and thus
$\dot{V}(t,{\bf x})$ is negative definite and the variance is globally asymptotically stabilizable.% \hfill$\blacksquare$

>>>>

\end{bmatrix} = \begin{bmatrix}
0 & 1 & \ldots & 0 & 0 \\
0 & 0 & \ldots& 0 & 0 \\
\vdots &  \vdots & \ddots & \vdots & \vdots \\
0 & 0  & \ldots & 0 & 1 \\
0 & 0 & \ldots& 0 & 0 
\end{bmatrix}  \begin{bmatrix}
x_{1}\\
\dot{x}_{1}\\
\vdots \\
x_{n}\\
\dot{x}_{n}
\end{bmatrix} + \begin{bmatrix}
0\\
1\\
\vdots\\
0\\
1
\end{bmatrix} u_x
\end{equation}
 However, just as with one robot, we can only control two states because the controllability matrix $\mathcal{C}_n$ has rank two:
\begin{equation}
\mathcal{C}_n=\left[ \begin{matrix} 
0\\
1\\
\vdots\\
0\\
1
\end{matrix}
\,\middle\vert\,
  \begin{matrix} 
1\\
0\\
\vdots\\
1\\
0
\end{matrix}
\,\middle\vert\,
\begin{matrix} 
0\\
0\\
\vdots\\
0\\
0
\end{matrix}
\,\middle\vert\,
\begin{matrix} 
0\\
0\\
\vdots\\
0\\
0
\end{matrix}\,\middle\vert\,
\ldots \right], \quad \textrm{rank}(\mathcal{C}_n)=2.
\end{equation}  
\subsection{Controlling the mean position}\label{sec:controlMeanPosition}
This means \emph{any} number of robots controlled by a global command have just two controllable states in each axis. We cannot arbitrarily control the position and velocity of two or more robots, but have options on which states to control.  %One option is to control the position and velocity of the $j^{th}$ robot. To find a potentially more useful option, 
We create the following reduced order system that represents the mean $x$ position and velocity of the $n$ robots:
\begin{align}
\begin{bmatrix}\nonumber
\dot{\bar{x}} \\
\ddot{\bar{x}}
\end{bmatrix} &= \frac{1}{n} \begin{bmatrix}
0& 1& \ldots &0& 1 \\
0& 0& \ldots &0& 0
\end{bmatrix}
\begin{bmatrix}
x_{1}\\
\dot{x}_{1}\\
\vdots\\
x_{n}\\
\dot{x}_{n}
\end{bmatrix} \\
&+ \frac{1}{n}\begin{bmatrix}
0& 0&  \ldots &0& 0 \\
0& 1&  \ldots &0& 1
\end{bmatrix}\begin{bmatrix} 
0\\
1\\
\vdots\\
0\\
1
\end{bmatrix} u_x
\end{align}
Thus:
\begin{equation}
\begin{bmatrix}
\dot{\bar{x}} \\
\ddot{\bar{x}}
\end{bmatrix} = \begin{bmatrix}
0& 1 \\
0& 0
\end{bmatrix}
\begin{bmatrix}
\bar{x}\\
\dot{\bar{x}}
\end{bmatrix} + \begin{bmatrix} 
0\\
1
\end{bmatrix} u_x
\end{equation}

@@

We will prove the origin is globally asymptotically stabilizable by using a control-Lyapunov function, as in \cite{Lyapunov1992}.  A suitable Lyapunov function is the squared variance error:
\begin{align}
\label{eq:LyapunovVariance}
V(t,\bf{x})  &= \frac{1}{2} (\sigma^2(\mathbf{x}) - \sigma^2_{goal})^2\nonumber\\
\dot{V}(t,\bf{x}) &= (\sigma^2(\mathbf{x})-\sigma^2_{goal})\dot{\sigma}^2(\mathbf{x})
\end{align}
We note here that $V(t,\mathbf{x})$ is positive definite and radially unbounded, and $V(t,\mathbf{x}) \equiv 0$ only at $\sigma^2(\mathbf{x}) = \sigma^2_{goal}$.
To make $\dot{V}(t,\mathbf{x})$ negative semi-definite, we choose
\begin{align}\label{eq:controlVariance}
u(t) &=   \begin{cases}
	 \mbox{move to wall} &\mbox{if } \sigma^2(\mathbf{x})>\sigma^2_{goal} \\ 
	 \mbox{move from wall} & \mbox{if } \sigma^2(\mathbf{x}) \le \sigma^2_{goal}.
\end{cases} 
\end{align}
 For such a $u(t)$,
 \begin{align}
\dot{\sigma}^2(\mathbf{x}) &=   \begin{cases}
	 \mbox{negative} &\mbox{if } \sigma^2(\mathbf{x})> \max(\sigma^2_{goal}, \sigma^2_{optimal}(n,r))  \\ 
	 W \epsilon & \mbox{if } \sigma^2(\mathbf{x}) \le \sigma^2_{goal},
\end{cases} 
\end{align} and thus
$\dot{V}(t,{\bf x})$ is negative definite and the variance is globally asymptotically stabilizable.% \hfill$\blacksquare$
